---
title: "Check Data Generating Function"
author: "Lacey Etzkorn"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/joint_simulate_data.R")
source("../delirium_package/R/random_weibull.R")
```


# Alpha 1, Alpha 2

```{r simulate}
if(!file.exists("../delirium_package_code/Explore_DataGenerating_Function.rdata")){
par0 = map2(as.list(rep(seq(-1, 1, length = 101),5)),
           as.list(rep(seq(-1, 1, length = 5),each = 101)),
           ~c(betaR = 1, etaR = 20,
         betaD = 1, etaD = 10,
         betaD2 = 1, etaD2 = 10,
         theta = 2, alpha1 = .x, alpha2 = .y,
         trtR = 0, trtD = 0, trtD2 = 0))
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 1200,truncate = 28,par0 = .))
save(data0, file = "../delirium_package_code/Explore_DataGenerating_Function.rdata")
}else  load(file = "../delirium_package_code/Explore_DataGenerating_Function.rdata")
```

# Plot Number of Events by Alpha1

For a fixed alpha2, a negative correlation between alpha1 and the total number of recurrent events.

```{r}
summary.df = 
tibble(alpha1 = rep(seq(-1, 1, length = 101),5),
       alpha2 = rep(seq(-1, 1, length = 5),each = 101),
       event.count = map(data0, ~summarise(., events = sum(event))) %>% unlist,
       t1 = map(data0, ~summarise(., events = sum(terminal1))) %>% unlist,
       t2 = map(data0, ~summarise(., events = sum(terminal1))) %>% unlist
       ) 
ggplot(summary.df,aes(x=alpha1, y = event.count, color = alpha2, group = alpha2))+
geom_point() +
stat_smooth()+
ylim(0,4000)
```


```{r}
# Plot Event Counts
ggplot(summary.df2,aes(x=alpha1, y = t1, color = alpha2, group = alpha2))+
geom_point() +
stat_smooth(se = F)
```

If either alpha1 or alpha2 are 1, then we see no effect of the other alpha on the total number of recurrent events.

This makes sense because if we expand the window over which we can observe terminal events by a certain factor (w), and we increase times between events by the same factor, we won't see a net increase in events.

But why can one terminal event have a strong positive association with the rate of recurrent events, and the other have a strong negative association with the rate of recurrent events, but

We expect there to be the most recurrent events when both alpha1 and alpha2 are -1, because this would mean the people with the most events also have long survival times. 


# Plot Event Counts by Total Time observed for each terminal event

```{r}
cordf <- 
map(data0,
    ~ group_by(., id) %>%
      summarise(n = sum(event),
                t = max(t),
                nt = n/t,
                type = ifelse(any(terminal1==1),1,
          	           ifelse(any(terminal2==1),2,0))) %>%
      summarise(cor1 = cor(nt[type==1], t[type==1]),
                cor2 = cor(nt[type==2], t[type==2]))
) %>%
do.call(what = "rbind")
cordf$alpha <- seq(-1,1, length = 1000)

ggplot(cordf) +
geom_point(aes(x = alpha, y = cor1)) +
stat_smooth(aes(x = alpha, y = cor1)) +
geom_point(aes(x = alpha, y = cor2), color = "red", alpha = 0.5) +
stat_smooth(aes(x = alpha, y = cor2)) +
geom_hline(yintercept = 0)
```

Even when the second terminal event is independent of the frailty (alpha2 = 0), 
we still see a correlation between time to terminal event 2 and the number of recurrent events among individuals 
who experienced terminal event 2.

If alpha1 = 1, and an individual survives a long time, but experiences event 2, that means they likely had a long time to event 1 (low frailty) and would have had fewer events than expected. Hence, we still see a negative correlation between time to event 2 and the number of recurrent events per unit time when alpha = 1.

Even when alpha2 = 0, there is still going to be a correlation between the two observed event times. (i.e. if one event time is long, the other will have been longer.)

```{r}
{
plot(cordf$cor1, cordf$cor2)
abline(h = 0)
abline(v = 0)
abline(a = c(0,1))
}
```

I think that when there is a relationship between the terminal event and rate of recurrent events,
it is hard to tell which terminal event is related to the recurrent event.



```{r}
ggplot(data0[[101]]) +
geom_density(aes(x = w, group = terminal1, fill = factor(terminal1)), alpha = 0.5)
```


# Decrease Theta

```{r simulate2}
if(!file.exists("../delirium_package_code/Explore_DataGenerating_Function2.rdata")){
par0 = map2(as.list(rep(seq(-1, 1, length = 101),5)),
           as.list(rep(seq(-1, 1, length = 5),each = 101)),
           ~c(betaR = 1, etaR = 5,
         betaD = 1, etaD = 10,
         betaD2 = 1, etaD2 = 10,
         theta = 0.5, alpha1 = .x, alpha2 = .y,
         trtR = 0, trtD = 0, trtD2 = 0))
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 1200,truncate = 28,par0 = .))
save(data0, file = "../delirium_package_code/Explore_DataGenerating_Function2.rdata")
}else  load(file = "../delirium_package_code/Explore_DataGenerating_Function2.rdata")

summary.df2 = 
tibble(alpha1 = rep(seq(-1, 1, length = 101),5),
       alpha2 = rep(seq(-1, 1, length = 5),each = 101),
       event.count = map(data0, ~summarise(., events = sum(event))) %>% unlist,
       t1 = map(data0, ~summarise(., events = sum(terminal1))) %>% unlist,
       t2 = map(data0, ~summarise(., events = sum(terminal1))) %>% unlist
       )

# Plot Event Counts
ggplot(summary.df2,aes(x=alpha1, y = event.count, color = alpha2, group = alpha2))+
geom_point() +
stat_smooth()
```

```{r}
# Plot Event Counts
ggplot(summary.df2,aes(x=alpha1, y = t1, color = alpha2, group = alpha2))+
geom_point() +
stat_smooth(se = F)
```
It does not look like the number of events changes with alpha1 and alpha2.

```{r}
summary.df2 %>% filter(alpha2==1) %>%
ggplot(aes(x=alpha1, y = event.count))+
geom_point() +
stat_smooth()
```



```{r}
library(mgcv)
mod1 <- 
gam(data = data, formula = g~trt + s(id, bs = "re"), 
    weights = data$event, #indicator for censoring
    family = cox.ph())

mod1 <- 
gamm(data = data, formula = g~trt, random = list(id = ~ 1),
     weights = data$event, #indicator for censoring
     family = cox.ph())

summary(mod1)
```
