---
title: "Simulation of Competing Joint Model"
output: html_document
---
```{r, echo =FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/competing_simulate_data.R")
source("../delirium_package/R/random_weibull.R")
source("../delirium_package/R/extract_frailtypack_coef_weibullgap.R")
```

We simulate an artificial data set where all baseline hazards are exponential (weibull shape = 1).
Simulated treatment effects are all 0.
We also set alpha1, alpha2 = 0 and theta = 0.5.

In the code below, I make the mistake of simulating from the gap time framework, 
but fitting the calendar times model.

# Competing Joint Model

$$
    r_{ij}(t|w_i, X_{ij}) = r_0(t) \exp(X_{i} \gamma_r + \frac 12 [w_i + u_i]) \\
    \lambda_{i}^{(1)}(t|w_i, X_{ij}) = \lambda_0^{(1)}(t) \exp(X_{i} \gamma_1 + \alpha_{1} w_i) \\
    \lambda_{i}^{(2)}(t|w_i, X_{ij}) = \lambda_0^{(2)}(t) \exp(X_{i} \gamma_2 + \alpha_{2} u_i) \\
        r_0(t) =  \frac{\eta_r \cdot t^{\eta_r-1}}{\beta_r^{\eta_r}} \text{(weibull hazard)}\\
    \lambda_0^{(1)}(t) = \frac{\beta_1 \cdot t^{\beta_1-1}}{\eta_1^{\beta_1}} \text{(weibull hazard)}\\
    \lambda_0^{(2)}(t) = \frac{\beta_2 \cdot t^{\beta_2-1}}{\eta_2^{\beta_2}} \text{(weibull hazard)}\\
    w_i \sim_{IID} N(0,\sigma^2)
$$


$$
    \textbf{Simulation Values:}\\
    \gamma_r = 0, \gamma_1 = 0, \gamma_2 = 0 \text{  (treatment effects)}\\
    \eta_r =20, \eta_1 = 10, \eta_2 = 10 \text{  (weibull scale)}\\
    \beta_r = 1, \beta_1 = 1, \beta_2 = 1 \text{  (weibull shape)}\\
    \alpha_1 = 0;  
    \alpha_2 = 0; 
    \theta = 0.5 \\
    \text{Maximum Observation (Censoring) Time } = 28
$$


# Test Shared Model 

```{r, fig.height = 14, fig.width=18}
set.seed(031821)
par00 = c(betaR = 1, etaR = 20,
         betaD = 1, etaD = 10,         
         betaD2 = 1, etaD2 = 10,
         theta = 0.5, alpha1 = 0, alpha2 = 0,
         trtR = 0, trtD = 0, trtD2 = 0)

data <- 
simulate.competing.data(n = 1200,
		truncate = 28,
		par0 = par00)
data %>% 
group_by(id) %>% 
summarise(del = any(event==1), 
          death = any(terminal1==1), 
          discharge = any(terminal2==1)) %>% 
summarise(mean(del), mean(death), mean(discharge), 
          mean(death*del)/mean(death), mean((discharge*del))/mean(discharge))
```


```{r}
mod0 <- 
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
jointGeneral = F,
initialize = FALSE, 
save.progress = T,
hazard = "Weibull",
gapTimes=F,
maxit = 100,
init.hazard = c(1,1,1,1,1,1),
init.Theta = 1,
init.Alpha1 = 0,
init.Alpha2 = 0,
init.B = c(0,0,0))

mod0$critCV
mod0$summary.table
```

# Simulate Data and Fit General Model

In this version, we will use the true data generating values as the starting point for the optimizer.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 8, fig.width = 12}
set.seed(012921)
nsim = 1000
par0 = rep(each = nsim, list(par00))

####################################################################
# Generate Data
if(!file.exists("../package_test_output/CompetingJointModels_20210723.rdata")){
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 1200,truncate = 2,par0 = .))

# Fit models
mod.joint.a <- list()

for(i in 1:length(par0)){
mod.joint.a[[i]] <-
multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal1) + terminal2(terminal2),
formula.terminalEvent = ~trt, 
formula.terminalEvent2= ~trt,
jointGeneral = F,
data = data0[[i]],
gapTimes=FALSE,
hazard = "Weibull",
maxit = 100, 
initialize = F,
init.hazard = par00[1:6],
init.Theta = par00[7],
init.Alpha1 = par00[8],
init.Alpha2 = par00[9],
init.B = c(0,0,0))
cat("Model ", i, "complete.\n")
save(data0, mod.joint.a, file = "../package_test_output/CompetingJointModels_20210723.rdata")
}
} else load(file = "../package_test_output/CompetingJointModels_20210723.rdata")
```


```{r, fig.height = 8, fig.width = 12}
####################################################################
# Combine  Results
truth =
tibble(Parameter = mod.joint.a[[1]]$summary.table$Parameter,
truth = par00)

results <- 
tibble(data = 1:(nsim),
       est = map(mod.joint.a, ~.$summary.table),
       error = map(mod.joint.a, ~.$critCV[2]) %>% unlist %>% 
           factor(labels = c("None","Calculation")),
       anyerror = map(mod.joint.a, ~.$critCV[2]!=1)%>% unlist,
       error2 = map(mod.joint.a, ~.$critCV[2] ==2) %>% unlist,
       error4 = map(mod.joint.a, ~.$critCV[2] ==4) %>% unlist) %>% 
unnest(c(est))

# Plot Results
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = data, x = Estimate, color = factor(error))) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "blue", "red"))+
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 1000, n = 500)")
```

```{r}
results %>% 
    filter(grepl("Alpha", Parameter)) %>%
    dplyr::select(data, Parameter, Raw) %>%
    pivot_wider(id_cols = data, names_from = "Parameter", values_from = "Raw") %>%
ggplot()+
    geom_point(aes(x = `Alpha, Terminal1`, y = `Alpha, Terminal2`))
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
tab_start <-
results %>%mutate(Truth = rep(truth$truth,nsim)) %>%
group_by(Parameter) %>%
summarise(Truth = Truth[1], Mean = mean(Estimate[!anyerror]),
          SD = sd(Estimate[!anyerror], na.rm=T),
          Est.SD = mean(Estimate.SE[!anyerror], na.rm=T), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value)

library(gt)
gt(tab_start, rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Joint Model Frailty Pack Fits (R = 1,000, n = 1,200)**"))%>%
cols_label(
    Truth = html("True Value"),
    Mean = html("Avg. Estimate"),
    SD = html("SE"),
    Est.SD = html("Avg. Est. SE")
)
```

# Errors

```{r}
mean(results$anyerror)
mean(results$error2)
mean(results$error4)
```
