---
title: "Simulation of Competing (Gap) Joint Model"
output: html_document
---
```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/competing_simulate_data.R")
source("../delirium_package/R/random_weibull.R")
source("../delirium_package/R/extract_frailtypack_coef_weibullgap.R")
```

In this simulation, we simulate from estimated values for the REDUCE gap times model, 
except theta = 0.4 instead of 5.4 and 
the scale parameters have been manipulated so that recurrent events happen more frequently (8 instead of 35)
and discharge is slightly prolonged (10 instead of 8.2).

# Competing Joint Model

$$
    r_{ij}(t|w_i, X_{ij}) = r_0(t) \exp(X_{i} \gamma_r + \frac 12 [w_i + u_i]) \\
    \lambda_{i}^{(1)}(t|w_i, X_{ij}) = \lambda_0^{(1)}(t) \exp(X_{i} \gamma_1 + \alpha_{1} w_i) \\
    \lambda_{i}^{(2)}(t|w_i, X_{ij}) = \lambda_0^{(2)}(t) \exp(X_{i} \gamma_2 + \alpha_{2} u_i) \\
        r_0(t) =  \frac{\eta_r \cdot t^{\eta_r-1}}{\beta_r^{\eta_r}} \text{(weibull hazard)}\\
    \lambda_0^{(1)}(t) = \frac{\beta_1 \cdot t^{\beta_1-1}}{\eta_1^{\beta_1}} \text{(weibull hazard)}\\
    \lambda_0^{(2)}(t) = \frac{\beta_2 \cdot t^{\beta_2-1}}{\eta_2^{\beta_2}} \text{(weibull hazard)}\\
    w_i \sim_{IID} N(0,\sigma^2)
$$
$$
    \textbf{Simulation Values:}\\
    \gamma_r = -0.08, \gamma_1 = 0.07, \gamma_2 = 0.08 \text{  (treatment effects)}\\
    \eta_r =8, \eta_1 = 25, \eta_2 = 10 \text{  (weibull scale)}\\
    \beta_r = 0.83, \beta_1 = 1.85, \beta_2 = 1.15 \text{  (weibull shape)}\\
    \alpha_1 = -0.12;  
    \alpha_2 = -0.32; 
    \theta = 0.4 \\
    \text{Maximum Observation (Censoring) Time} = 28
$$

# Test Shared Model 

```{r, fig.height = 14, fig.width=18}
set.seed(080721)
par00 = c(betaR = 1, etaR = 10,
         betaD = 1.85, etaD = 25,      
         betaD2 = 1.15, etaD2 = 10,
         theta = 0.2, alpha1 = -0.12, alpha2 = -0.32,
         trtR = -0.08, trtD = 0.07, trtD2 = 0.08)

# put in treatment effects
# opposite directions for alpha
# alpha1 = 0.15
# alpha2 = -0.3

data <- 
simulate.competing.data(n = 1500, truncate = 28, par0 = par00)

data %>% 
group_by(id) %>% 
summarise(del = any(event==1),
          no.del = sum(event==1),
          death = any(terminal1==1), 
          discharge = any(terminal2==1)) %>% 
summarise(mean(del), max(no.del), mean(death), mean(discharge), 
          mean(death*del)/mean(death), mean((discharge*del))/mean(discharge)) %>% 
t
```


```{r}
mod0 <- 
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
jointGeneral = F,
initialize = T, 
save.progress = T,
hazard = "Weibull",
gapTimes=T,
maxit = 350)

mod0b <- # without initialization
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
jointGeneral = F,
initialize = F, 
save.progress = T,
hazard = "Weibull",
gapTimes=T,
maxit = 350)

mod0$critCV
mod0b$critCV
```

```{r}
moda <- 
frailtyPenal(formula = Surv(g, event) ~ trt + cluster(id) +
		terminal(terminal1),
formula.terminalEvent = ~trt, 
jointGeneral = F,
data = data,
recurrentAG=F,
hazard = "Weibull",RandDist = "LogN",
maxit = 200, init.Theta = .4)
moda$istop
```

# Simulate Data and Fit Shared Competing Model

In this version, we will use the true data generating values as the starting point for the optimizer.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 8, fig.width = 12}
set.seed(210806)
nsim = 500
par0 = rep(each = nsim, list(par00))

####################################################################
# Generate Data
if(!file.exists("../package_test_output/CompetingJointModels_20210807.rdata")){
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 1200,truncate = 2,par0 = .))

# Fit models
mod.joint.a <- list()
mod.joint.b <- list()
mod.joint.c <- list()
mod.joint.d <- list()

for(i in 1:length(par0)){
mod.joint.a[[i]] <-
multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal1) + terminal2(terminal2),
formula.terminalEvent = ~trt, 
formula.terminalEvent2= ~trt,
jointGeneral = F,
data = data0[[i]],
gapTimes=T,
hazard = "Weibull",
maxit = 100, 
initialize = F,
init.hazard = par00[1:6],
init.Theta = par00[7],
init.Alpha1 = par00[8],
init.Alpha2 = par00[9],
init.B = c(0,0,0))

### Death Only
mod.joint.b[[i]] <-
frailtyPenal(formula = Surv(g, event) ~ trt + cluster(id) +
		terminal(terminal1) ,
formula.terminalEvent = ~trt, 
jointGeneral = F,
data = data0[[i]],
recurrentAG=F,
hazard = "Weibull",RandDist = "LogN",
maxit = 100)

### Discharge Only
mod.joint.c[[i]] <-
frailtyPenal(formula = Surv(g, event) ~ trt + cluster(id) +
		terminal(terminal2) ,
formula.terminalEvent = ~trt, 
jointGeneral = F,
data = data0[[i]],
recurrentAG=F,
hazard = "Weibull",RandDist = "LogN",
maxit = 100)

### Death+Discharge
data0[[i]]$terminal <- data0[[i]]$terminal1 + data0[[i]]$terminal2 
mod.joint.d[[i]] <-
frailtyPenal(formula = Surv(g, event) ~ trt + cluster(id) +
		terminal(terminal) ,
formula.terminalEvent = ~trt, 
jointGeneral = F,
data = data0[[i]],
recurrentAG=F,
hazard = "Weibull",RandDist = "LogN",
maxit = 100)

cat("Model ", i, "complete.\n")
save(data0, mod.joint.a, mod.joint.b, mod.joint.c,mod.joint.d, 
     file = "../package_test_output/CompetingJointModels_20210807.rdata")
}
} else load(file = "../package_test_output/CompetingJointModels_20210807.rdata")
```

# Verify Competing Joint Model Works

```{r, fig.height = 8, fig.width = 12}
####################################################################
# Combine  Results
truth =
tibble(Parameter = mod.joint.a[[1]]$summary.table$Parameter,
       truth = par00)

results <- 
tibble(data = 1:(nsim),
       est = map(mod.joint.a, ~.$summary.table),
       error = map(mod.joint.a, ~.$critCV[2]) %>% unlist %>% 
           factor(labels = c("None","Maxit","Calculation")),
       anyerror = map(mod.joint.a, ~.$critCV[2]!=1)%>% unlist,
       error2 = map(mod.joint.a, ~.$critCV[2] ==2) %>% unlist,
       error4 = map(mod.joint.a, ~.$critCV[2] ==4) %>% unlist) %>% 
unnest(c(est))

# Plot Results
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = data, x = Estimate, color = factor(error))) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "blue", "red"))+
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 1000, n = 1200)")
```

```{r}
results %>% 
    filter(grepl("Alpha", Parameter)) %>%
    dplyr::select(data, Parameter, Raw) %>%
    pivot_wider(id_cols = data, names_from = "Parameter", values_from = "Raw") %>%
ggplot()+
    geom_point(aes(x = `Alpha, Terminal1`, y = `Alpha, Terminal2`))
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
tab_start <-
results %>%mutate(Truth = rep(truth$truth,nsim)) %>%
group_by(Parameter) %>%
summarise(Truth = Truth[1], 
          Mean = median(Estimate[!anyerror], trim = 0.005),
          SD = mean((Estimate[!anyerror]-Mean)^2, na.rm=T, trim = 0.005)^.5,
          Est.SD = mean(Estimate.SE[!anyerror], na.rm=T, trim = 0.005), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value)

library(gt)
gt(tab_start, rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Joint Model Frailty Pack Fits (R = 1,000, n = 1,200)**"))%>%
cols_label(
    Truth = html("True Value"),
    Mean = html("Avg. Estimate"),
    SD = html("SE"),
    Est.SD = html("Avg. Est. SE")
)
```

## Errors

```{r}
mean(results$anyerror)
mean(results$error2)
mean(results$error4)
```

# Comparison with Joint Model

```{r}
Parameter = mod.joint.a[[1]]$summary.table$Parameter
results <- list()
for(i in 1:nsim){
results[[i]]<- 
tibble(sim = i,
       Parameter = c(Parameter[c(1:4,7:8,10:11)], #model1a
	         Parameter[c(1:2,5:6,7,9,10,12)],
	         Parameter),
       Estimates = c(mod.joint.b[[i]]$b[1:5]^2, mod.joint.b[[i]]$b[6:8], 
       	         mod.joint.c[[i]]$b[1:5]^2, mod.joint.c[[i]]$b[6:8],
       	         mod.joint.a[[i]]$summary.table$Estimate) %>% round(3),
       SE = c(sqrt(diag(mod.joint.b[[i]]$varHtotal))[1:5]*2*sqrt(mod.joint.b[[i]]$b[1:5]), 
       	  sqrt(diag(mod.joint.b[[i]]$varHtotal))[6:8], 
       	  sqrt(diag(mod.joint.c[[i]]$varHtotal))[1:5]*2*sqrt(mod.joint.c[[i]]$b[1:5]),
       	  sqrt(diag(mod.joint.c[[i]]$varHtotal))[6:8],
       	  mod.joint.a[[i]]$summary.table$Estimate.SE) %>% round(3),
       Error = c(rep(mod.joint.b[[1]]$istop != 1, 8),
                 rep(mod.joint.c[[1]]$istop != 1, 8),
                 rep(mod.joint.a[[1]]$critCV[2] != 1, 12))) %>%
mutate(model = c(rep("Death",8), 
       	         rep("Discharge",8), 
       	         rep("Competing",12)))
}

results %>%
do.call(what = "rbind") %>% 
group_by(Parameter, model) %>%
summarise(Mean = median(Estimates[!Error], trim = 0.005),
          SD = sd(Estimates[!Error]),
          Est.SD = median(SE[!Error], na.rm=T), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value) %>%
pivot_wider(id_cols = Parameter, names_from = c(model), values_from = c(Mean, SD, Est.SD) ) %>%
#arrange(Parameter)  %>%
relocate(SD_Death, .after = Mean_Death)%>%
relocate(SD_Discharge, .after = Mean_Discharge)%>%
relocate(SD_Competing, .after = Mean_Competing)%>%
relocate(SD_Death, .after = Mean_Death)%>%
relocate(SD_Discharge, .after = Mean_Discharge)%>%
relocate(SD_Competing, .after = Mean_Competing) %>%
dplyr::select(-Est.SD_Competing:-Est.SD_Discharge) %>%
arrange(c(5:6,1:3,7:12,4))%>%
mutate(Truth = par00[c(10:12,7:9,2,4,6,1,3,5)]) %>%
gt(rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Simulation Results: Weibull Model Estimates**"))%>%
cols_label(
    Mean_Death = html("Joint: Death"),
    Mean_Discharge = html("Joint: Discharge"),
    Mean_Competing = html("Competing"),
    Mean_Death = html("Joint (Death)"),
    Mean_Discharge = html("Joint (Discharge)"),
    Mean_Competing = html("Competing"),
    SD_Death = html("(SE)"),
    SD_Discharge = html("(SE)"),
    SD_Competing = html("(SE)"),
)%>%
fmt_missing(columns = 1:8, missing_text = "")  %>%
 tab_style(
     style = list(
       cell_borders(sides = "left")
     ),
     locations = list(
       cells_body(
         columns = vars(Truth)
       )
     )
   ) %>% 
text_transform(
  locations = cells_body(columns = c(3,5,7)),
  fn = function(x) ifelse(x == "","",paste0("(",x,")"))
  )%>%
cols_align(
  align = "left",
  columns = c(3,5,7)
)
```
