---
title: "Simulation of Gap Time Competing Joint Model (Gap Times)"
output: html_document
---
```{r, echo =FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/competing_simulate_data.R")
source("../delirium_package/R/random_weibull.R")
source("../delirium_package/R/extract_frailtypack_coef_weibullgap.R")
```


Simulate gap times data based on REDUCE trial estimates for calendar times. 
In the code below, I make the mistake of simulating from the gap time framework, 
but fitting the calendar times model.

# Competing Joint Model

$$
    r_{ij}(t|w_i, X_{ij}) = r_0(t) \exp(X_{i} \gamma_r + w_i) \\
    \lambda_{i}^{(1)}(t|w_i, X_{ij}) = \lambda_0^{(1)}(t) \exp(X_{i} \gamma_1 + \alpha_{1} w_i) \\
    \lambda_{i}^{(2)}(t|w_i, X_{ij}) = \lambda_0^{(2)}(t) \exp(X_{i} \gamma_2 + \alpha_{2} w_i) \\
        r_0(t) =  \frac{\eta_r \cdot t^{\eta_r-1}}{\beta_r^{\eta_r}} \text{(weibull hazard)}\\
    \lambda_0^{(1)}(t) = \frac{\beta_1 \cdot t^{\beta_1-1}}{\eta_1^{\beta_1}} \text{(weibull hazard)}\\
    \lambda_0^{(2)}(t) = \frac{\beta_2 \cdot t^{\beta_2-1}}{\eta_2^{\beta_2}} \text{(weibull hazard)}\\
    w_i \sim_{IID} N(0,\sigma^2)
$$

$$
    \textbf{Simulation Values: (REDUCE Calendar Time Estimates from 7/21/21)}\\
    \gamma_r = -0.041 , \gamma_1 = 0.115 , \gamma_2 = 0.172 \text{  (treatment effects)}\\
    \eta_r = 87.62, \eta_1 = 18.48, \eta_2 = 8.4 \text{  (weibull scale)}\\
    \beta_r = 0.40, \beta_1 = 2.13, \beta_2 = 2.06 \text{  (weibull shape)}\\
    \alpha_1 = -0.88;  
    \alpha_2 = -2.9; 
    \theta = 0.44 \\
    \text{Maximum Observation (Censoring) Time } = 28
$$

# Test Shared Model 

```{r, fig.height = 14, fig.width=18}
set.seed(031821)
data <- 
simulate.competing.data(n = 1200,
		truncate = 28,
		par0 = c(betaR = 0.40, etaR = 87.62,
         betaD = 2.13, etaD = 18.48,         
         betaD2 = 2.06, etaD2 = 8.4,
         theta = 0.44, alpha1 = -0.88, alpha2 = -2.9,
         trtR = -0.041, trtD = 0.115, trtD2 = 0.172))

mod0 <- 
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
jointGeneral = F,
initialize = FALSE, 
save.progress = T,
hazard = "Weibull",
gapTimes=F,
maxit = 100,
init.hazard = c(1,1,1,1,1,1),
init.Theta = 1,
init.Alpha1 = 0,
init.Alpha2 = 0,
init.B = c(0,0,0))

mod0$critCV
mod0$summary.table
```

# Simulate Data and Fit General Model

In this version, we will use the true data generating values as the starting point for the optimizer.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 8, fig.width = 12}
set.seed(012921)
nsim = 1000
par0 = rep(each = nsim, list(
       c(betaR = 0.40, etaR = 87.62,
         betaD = 2.13, etaD = 18.48,         
         betaD2 = 2.06, etaD2 = 8.4,
         theta = 0.44, alpha1 = -0.88, alpha2 = -2.9,
         trtR = -0.041, trtD = 0.115, trtD2 = 0.172)))

####################################################################
# Generate Data
if(!file.exists("../package_test_output/CompetingJointModels_20210721.rdata")){
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 500,truncate = 2,par0 = .))

# Fit models
mod.joint.a <- list()

for(i in 1:length(par0)){
mod.joint.a[[i]] <-
multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal1) + terminal2(terminal2),
formula.terminalEvent = ~trt, 
formula.terminalEvent2= ~trt,
jointGeneral = F,
data = data0[[i]],
gapTimes=FALSE,
hazard = "Weibull",
maxit = 100, 
initialize = F,
init.hazard = c(1,50,1,20,1,20),
init.Theta = 0.5,
init.Alpha1 = 0,
init.Alpha2 = 0,
init.B = c(0,0,0))
cat("Model ", i, "complete.\n")
save(data0, mod.joint.a, file = "../package_test_output/CompetingJointModels_20210721.rdata")
}
} else load(file = "../package_test_output/CompetingJointModels_20210721.rdata")
```




```{r, fig.height = 8, fig.width = 12}
####################################################################
# Combine  Results
truth =
tibble(Parameter = mod.joint.a[[1]]$summary.table$Parameter,
truth = c(betaR = 0.40, etaR = 87.62,
         betaD = 2.13, etaD = 18.48,         
         betaD2 = 2.06, etaD2 = 8.4,
         theta = 0.44, alpha1 = -0.88, alpha2 = -2.9,
         trtR = -0.041, trtD = 0.115, trtD2 = 0.172)
       )

results <- 
tibble(data = 1:(nsim),
       est = map(mod.joint.a, ~.$summary.table),
       error = map(mod.joint.a, ~.$critCV[2]) %>% unlist %>% 
           factor(labels = c("None","Max Iter", "Calculation")),
       anyerror = map(mod.joint.a, ~.$critCV[2]!=1)%>% unlist,
       error2 = map(mod.joint.a, ~.$critCV[2] ==2) %>% unlist,
       error4 = map(mod.joint.a, ~.$critCV[2] ==4) %>% unlist) %>% 
unnest(c(est))

# Plot Results
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = data, x = Estimate, color = factor(error))) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "blue", "red"))+
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 1000, n = 500)")
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
tab_start <-
results %>%mutate(Truth = rep(truth$truth,nsim)) %>%
group_by(Parameter) %>%
summarise(Truth = Truth[1], Mean = mean(Estimate[!anyerror]),
          SD = sd(Estimate[!anyerror], na.rm=T),
          Est.SD = mean(Estimate.SE[!anyerror], na.rm=T), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value)

library(gt)
gt(tab_start, rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Joint Model Frailty Pack Fits (R = 1,000, n = 1,200)**"))%>%
cols_label(
    Truth = html("True Value"),
    Mean = html("Avg. Estimate"),
    SD = html("SE"),
    Est.SD = html("Avg. Est. SE")
)
```

# Errors

```{r}
mean(results$anyerror)
mean(results$error2)
mean(results$error4)
```

# Identifiability Conditions for the General Model

If $\alpha_1 , \alpha_2 = 0$, then, essentially, the model consists of only one random effect $u_i + w_i$, so $u_i$ and $w_i$ are not distinguishable, and separate variance components cannot be estimated.

If one of $\alpha_1$ or $\alpha_2$ are 0, then we might also have some identifiability problems.
