---
title: "Simulation of Competing (Gap) Joint Model"
output: html_document
---
```{r, echo =FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/competing_simulate_data.R")
source("../delirium_package/R/random_weibull.R")
source("../delirium_package/R/extract_frailtypack_coef_weibullgap.R")
```

No treatment effects, but what if alpha1 and alpha2 are in the same direction.

We simulate an artificial data set where shape parameters match estimates from the reduce trial.
Scale parameters are chosen by us (20,25,10).
Simulated treatment effects are all 0.
We also set alpha1 = -0.3, alpha2 = -0.1 and theta = 1.

In the code below, I make the mistake of simulating from the gap time framework, 
but fitting the calendar times model.

# Competing Joint Model

$$
    r_{ij}(t|w_i, X_{ij}) = r_0(t) \exp(X_{i} \gamma_r + \frac 12 [w_i + u_i]) \\
    \lambda_{i}^{(1)}(t|w_i, X_{ij}) = \lambda_0^{(1)}(t) \exp(X_{i} \gamma_1 + \alpha_{1} w_i) \\
    \lambda_{i}^{(2)}(t|w_i, X_{ij}) = \lambda_0^{(2)}(t) \exp(X_{i} \gamma_2 + \alpha_{2} u_i) \\
        r_0(t) =  \frac{\eta_r \cdot t^{\eta_r-1}}{\beta_r^{\eta_r}} \text{(weibull hazard)}\\
    \lambda_0^{(1)}(t) = \frac{\beta_1 \cdot t^{\beta_1-1}}{\eta_1^{\beta_1}} \text{(weibull hazard)}\\
    \lambda_0^{(2)}(t) = \frac{\beta_2 \cdot t^{\beta_2-1}}{\eta_2^{\beta_2}} \text{(weibull hazard)}\\
    w_i \sim_{IID} N(0,\sigma^2)
$$

Hence, we can generate data from the shared model using fits from the REDUCE trial.

$$
    \textbf{Simulation Values:}\\
    \gamma_r = 0, \gamma_1 = 0, \gamma_2 = 0 \text{  (treatment effects)}\\
    \eta_r =20, \eta_1 = 25, \eta_2 = 10 \text{  (weibull scale)}\\
    \beta_r = 1, \beta_1 = 1.8, \beta_2 = 1.15 \text{  (weibull shape)}\\
    \alpha_1 = -0.3;  
    \alpha_2 = -0.1; 
    \theta = 1 \\
    \text{Maximum Observation (Censoring) Time } = 28
$$

# Test Competing Joint Model 

```{r, fig.height = 14, fig.width=18}
set.seed(080421)
par00 = c(betaR = 1, etaR = 20,
         betaD = 1.8, etaD = 25,         
         betaD2 = 1.15, etaD2 = 10,
         theta = 1, alpha1 = -0.3, alpha2 = -0.1,
         trtR = 0, trtD = 0, trtD2 = 0)

data <- 
simulate.competing.data(n = 1200,
		truncate = 28,
		par0 = par00)

data %>% 
group_by(id) %>% 
summarise(del = any(event==1), 
          death = any(terminal1==1), 
          discharge = any(terminal2==1)) %>% 
summarise(mean(del), mean(death), mean(discharge), 
          mean(death*del)/mean(death), mean((discharge*del))/mean(discharge)) %>% 
t
```


```{r}
mod0 <- 
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
jointGeneral = F,
initialize = FALSE, 
save.progress = T,
hazard = "Weibull",
gapTimes=F,
maxit = 200,
init.hazard = c(1,1,1,1,1,1),
init.Theta = 1,
init.Alpha1 = -0.25,
init.Alpha2 = -0.1,
init.B = c(0,0,0))

mod0$critCV
mod0$summary.table
```

# Simulate Data and Fit General Model

In this version, we will use the true data generating values as the starting point for the optimizer.

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 8, fig.width = 12}
set.seed(012921)
nsim = 1000
par0 = rep(each = nsim, list(par00))

####################################################################
# Generate Data
if(!file.exists("../package_test_output/CompetingJointModels_20210803.rdata")){
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 1200,truncate = 2,par0 = .))

# Fit models
mod.joint.a <- list()

for(i in 1:length(par0)){
mod.joint.a[[i]] <-
multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal1) + terminal2(terminal2),
formula.terminalEvent = ~trt, 
formula.terminalEvent2= ~trt,
jointGeneral = F,
data = data0[[i]],
gapTimes=FALSE,
hazard = "Weibull",
maxit = 100, 
initialize = F,
init.hazard = par00[1:6],
init.Theta = par00[7],
init.Alpha1 = par00[8],
init.Alpha2 = par00[9],
init.B = c(0,0,0))
cat("Model ", i, "complete.\n")
save(data0, mod.joint.a, file = "../package_test_output/CompetingJointModels_20210803.rdata")
}
} else load(file = "../package_test_output/CompetingJointModels_20210803.rdata")

####################################################################
# Combine  Results
truth =
tibble(Parameter = mod.joint.a[[1]]$summary.table$Parameter,
truth = par00)

results <- 
tibble(data = 1:(nsim),
       est = map(mod.joint.a, ~.$summary.table),
       error = map(mod.joint.a, ~.$critCV[2]) %>% unlist %>% 
           factor(labels = c("None","Maxit","Calculation")),
       anyerror = map(mod.joint.a, ~.$critCV[2]!=1)%>% unlist,
       error2 = map(mod.joint.a, ~.$critCV[2] ==2) %>% unlist,
       error4 = map(mod.joint.a, ~.$critCV[2] ==4) %>% unlist) %>% 
unnest(c(est))

# Plot Results
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = data, x = Estimate, color = factor(error))) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "blue", "red"))+
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 1000, n = 1200)")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
tab_start <-
results %>%mutate(Truth = rep(truth$truth,nsim)) %>%
group_by(Parameter) %>%
summarise(Truth = Truth[1], Mean = mean(Estimate[!anyerror]),
          SD = sd(Estimate[!anyerror], na.rm=T),
          Est.SD = mean(Estimate.SE[!anyerror], na.rm=T), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value)

library(gt)
gt(tab_start, rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Joint Model Frailty Pack Fits (R = 1,000, n = 1,200)**"))%>%
cols_label(
    Truth = html("True Value"),
    Mean = html("Avg. Estimate"),
    SD = html("SE"),
    Est.SD = html("Avg. Est. SE")
)
```

# Errors

```{r}
mean(results$anyerror)
mean(results$error2)
mean(results$error4)
```
