---
title: "FrailtyPenal Tests"
output:
  html_document:
    df_print: paged
---

# Generate the Data Sets

I have already run this section for you to produce the list of data sets `data0`.

```{r setup, eval=FALSE}
source("../frailtypack/R/simulate_competing_data.R")
source("../delirium_package/R/random_weibull.R")

set.seed(10)
# Simulate 10 data sets and store them into a list
data0 <- 
map(1:10, ~simulate.competing.data(n = 500,
		         par0 = c(betaR = .8, etaR = 25,
		                  betaD = 1, etaD = 84,
		                  betaD2 = 1, etaD2 = 84,
		                  theta = 1,
		                  alpha1 = 1.4, alpha2 = 1.4,
		                  trtR = -.3, trtD = -.5, trtD2 = -.5)) %>%
    	mutate(terminal = terminal1 + terminal2,
    	       g10 = g/10,
    	       g5 = g/5,
    	       g20 = g/20,
    	       g.2 = g*2))
save(data0, file = "FrailtyPenal_Test_Data.rdata")
```

# Verify Data

Verify that no one has a recurrent event at the final observation time.

```{r}
library(frailtypack); library(tidyverse)
load("FrailtyPenal_Test_Data.rdata")
tibble(iter = 1:10, data = data0) %>% 
unnest(data)%>%
group_by(iter, id) %>% 
filter(t == max(t)) %>%
ungroup() %>%
summarise(any(event == 1))
```

Verify max of one terminal event per person

```{r}
tibble(iter = 1:10, data = data0) %>% 
unnest(data)%>%
group_by(iter, id) %>% 
summarise(n = sum(terminal)) %>%
ungroup() %>%
summarise(all(n <= 1))
```

Verify all gap times are positive.

```{r}
tibble(iter = 1:10, data = data0) %>% 
unnest(data)%>%
summarise(all(g > 0))
```

# Fit Joint Models: Weibull-Gamma Gap Time Models

### Original Time Scale

```{r fit_models}
weibull.gamma.gap.models <-
map(data0, 
~frailtyPenal(formula = Surv(g, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=FALSE,
	hazard = "Weibull",
	maxit = 100)
)
```

### Gap Times/10

```{r }
weibull.gamma.gap10.models <-
map(data0, 
~frailtyPenal(formula = Surv(g10, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=FALSE,
	hazard = "Weibull",
	maxit = 100)
)
```

### Gap Times/5

```{r }
weibull.gamma.gap5.models <-
map(data0, 
~frailtyPenal(formula = Surv(g5, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=FALSE,
	hazard = "Weibull",
	maxit = 100)
)
```

### Gap Times/20

```{r }
weibull.gamma.gap20.models <-
map(data0, 
~frailtyPenal(formula = Surv(g20, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=FALSE,
	hazard = "Weibull",
	maxit = 100)
)
```

### Gap Times * 2

```{r }
weibull.gamma.gap.2.models <-
map(data0, 
~frailtyPenal(formula = Surv(g.2, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=FALSE,
	hazard = "Weibull",
	maxit = 100)
)
```

## Error Codes

```{r errors}
error.df <- 
tibble(
dataset = 1:10,
orig = map(weibull.gamma.gap.models, ~.$istop) %>% unlist,
fifth = map(weibull.gamma.gap5.models, ~.$istop) %>% unlist,
tenth = map(weibull.gamma.gap10.models, ~.$istop) %>% unlist,
twentyth = map(weibull.gamma.gap20.models, ~.$istop) %>% unlist,
double = map(weibull.gamma.gap.2.models, ~.$istop) %>% unlist
)
error.df
```

#### Number Models Converged by Timescale

```{r}
error.df[,-1] %>% lapply(function(x) sum(x == 1)) %>% unlist
```

# Fit Joint Models: Weibull-Gamma Calendar Time Models

```{r }
weibull.gamma.cal.models <-
map(data0, 
~frailtyPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=T, # changed to true for calendar model
	hazard = "Weibull",
	maxit = 100)
)
```

```{r}
map(weibull.gamma.cal.models, ~.$istop) %>% unlist
```

# Fit Joint Models: Weibull-LogNormal Gap Time Models

```{r }
weibull.logn.gap.models <-
map(data0, 
~frailtyPenal(formula = Surv(g, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=FALSE,
	hazard = "Weibull",
	RandDist = "LogN",
	maxit = 100)
)
```


```{r}
map(weibull.logn.gap.models, ~.$istop) %>% unlist
```

# Fit Joint Models: Weibull-LogNormal Calendar Time Models

```{r }
weibull.logn.cal.models <-
map(data0, 
~frailtyPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal),
	formula.terminalEvent = ~trt,
	data = .,
	recurrentAG=T,
	hazard = "Weibull",
	RandDist = "LogN",
	maxit = 100)
)
```

```{r}
map(weibull.logn.cal.models, ~.$istop) %>% unlist
```

```{r}
save(weibull.gamma.gap.models,
     weibull.gamma.gap5.models, 
     weibull.gamma.gap10.models,
     weibull.gamma.gap20.models,
     weibull.gamma.gap.2.models,
     weibull.gamma.cal.models,
     weibull.logn.gap.models,
     weibull.logn.cal.models,
     file = "Two_Terminal_Event_Models.rdata"
)
```
