---
title: "Simulation of Competing (Gap) Joint Model"
output: html_document
---
```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/competing_simulate_data.R")
source("../delirium_package/R/random_weibull.R")
save.date <- "20211213"
set.seed(121321)
```

In this simulation, we simulate from estimated values for the REDUCE gap times model, 
except theta = 0.4 instead of 5.4 and 
the scale parameters have been manipulated so that recurrent events happen more frequently (8 instead of 35)
and discharge is slightly prolonged (10 instead of 8.2).

# Competing Joint Model

$$
    r_{ij}(t|w_i, X_{ij}) = r_0(t) \exp(X_{i} \gamma_r + \frac 12 [w_i + u_i]) \\
    \lambda_{i}^{(1)}(t|w_i, X_{ij}) = \lambda_0^{(1)}(t) \exp(X_{i} \gamma_1 + \alpha_{1} w_i) \\
    \lambda_{i}^{(2)}(t|w_i, X_{ij}) = \lambda_0^{(2)}(t) \exp(X_{i} \gamma_2 + \alpha_{2} u_i) \\
        r_0(t) =  \frac{\eta_r \cdot t^{\eta_r-1}}{\beta_r^{\eta_r}} \text{(weibull hazard)}\\
    \lambda_0^{(1)}(t) = \frac{\beta_1 \cdot t^{\beta_1-1}}{\eta_1^{\beta_1}} \text{(weibull hazard)}\\
    \lambda_0^{(2)}(t) = \frac{\beta_2 \cdot t^{\beta_2-1}}{\eta_2^{\beta_2}} \text{(weibull hazard)}\\
    w_i \sim_{IID} N(0,\sigma^2)
$$
# Simulation Values

```{r, fig.height = 14, fig.width=18}
par00 = c(betaR = 1, etaR = 10,
         betaD = 1.85, etaD = 25,      
         betaD2 = 1.15, etaD2 = 10,
         theta = 0.2, alpha1 = -0.12, alpha2 = -0.32,
         trtR = -0.08, trtD = 0.07, trtD2 = 0.08)
print(par00)
```

# Simulate One Data Set

```{r, fig.height = 14, fig.width=18}
data <- 
simulate.competing.data(n = 1500, truncate = 28, par0 = par00)

data %>% 
group_by(id) %>% 
summarise(del = any(event==1),
          no.del = sum(event==1),
          death = any(terminal1==1), 
          discharge = any(terminal2==1)) %>% 
summarise(`Delirium Prevalence` = mean(del), 
          `Max No. Delirium Events` = max(no.del), 
          `Proportion Died` = mean(death), 
          `Proportion Discharged` = mean(discharge), 
          `Delirium Prevalence among Mortalities` = mean(death*del)/mean(death), 
          `Delirium Prevalence among Discharges` = mean((discharge*del))/mean(discharge)) %>% 
t
```

# Fit One Competing Model

```{r}
mod0 <- 
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
jointGeneral = F,
initialize = T, 
save.progress = T,
hazard = "Weibull",
gapTimes=T,
maxit = 350)

mod0$summary.table
```

# Simulate Many Data Sets and Fit Competing Models

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 8, fig.width = 12}
nsim = 10
par0 = rep(each = nsim, list(par00))

####################################################################
# Generate Data
if(!file.exists(paste0("../package_test_output/CompetingJointModels_",save.date,".rdata"))){
	data0 <- 
	map(par0, 
	    ~simulate.competing.data(n = 1200,truncate = 2,par0 = .))
	mod.joint <- list()
	# Fit models
	for(i in 1:length(par0)){
		mod.joint[[i]] <-
		multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
				terminal(terminal1) + terminal2(terminal2),
			formula.terminalEvent = ~trt, 
			formula.terminalEvent2= ~trt,
			jointGeneral = F,
			save.progress = T,
			data = data0[[i]],
			gapTimes=T,
			hazard = "Weibull",
			maxit = 100, 
			initialize = T)
		cat("Model ", i, "complete.\n")
		save(data0, mod.joint, 
		     file = paste0("../package_test_output/CompetingJointModels_",save.date,".rdata"))
	}
} else load(file = paste0("../package_test_output/CompetingJointModels_",save.date,".rdata"))
```

# Verify Competing Joint Model Estimates

```{r, fig.height = 8, fig.width = 12}
truth =
tibble(Parameter = mod.joint[[1]]$summary.table$Parameter,
       truth = par00)

results <- 
tibble(data = 1:(nsim),
       est = map(mod.joint, ~.$summary.table),
       error = map(mod.joint, ~.$critCV[2]) %>% unlist %>% 
           factor(labels = c("None","Maxit","Calculation"),levels = c(1,2,4)),
       anyerror = map(mod.joint, ~.$critCV[2]!=1)%>% unlist,
       error2 = map(mod.joint, ~.$critCV[2] ==2) %>% unlist,
       error4 = map(mod.joint, ~.$critCV[2] ==4) %>% unlist) %>% 
unnest(c(est))

# Plot Results
results %>% 
ggplot()+
geom_point(aes(y = data, x = Estimate, color = error)) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "green", "red"))+
theme_bw()+
labs(title = paste0("Competing Joint Model Simulation Results(R =",length(par00),", n = ", length(unique(data0[[1]]$id)),")"))
```

```{r}
truth =
tibble(Parameter = mod.joint[[1]]$summary.table$Parameter,
       truth = par00)

results2 <- 
tibble(data = 1:(nsim),
       par1 = map(mod.joint, ~.$initialization$summary.table1$Parameter),
       est1 = map(mod.joint, ~.$initialization$summary.table1$Estimate),
       par2 = map(mod.joint, ~.$initialization$summary.table2$Parameter),
       est2 = map(mod.joint, ~.$initialization$summary.table2$Estimate)) %>% 
unnest(c(par1, est1, par2, est2))

# Plot Results
results2 %>%
filter(par1 %in% c("Shape, Recurrent","Scale, Recurrent", "Sigma, Terminal", "Beta1")) %>%
ggplot()+
geom_point(aes(y = data, x = est1)) +
facet_wrap("par1", scales = "free") +
geom_vline(data = tibble(truth = par00[c("betaR","etaR","theta","trtR")],
		 par1 = c("Shape, Recurrent","Scale, Recurrent", "Sigma, Terminal", "Beta1")),
           aes(xintercept = truth), color = "red") +
theme_bw()+
labs(title = paste0("Competing Joint Model Simulation Results(R =",length(par00),", n = ", length(unique(data0[[1]]$id)),")"))
```


```{r}
results %>% 
    filter(grepl("Alpha", Parameter)) %>%
    dplyr::select(data, Parameter, Raw) %>%
    pivot_wider(id_cols = data, names_from = "Parameter", values_from = "Raw") %>%
ggplot()+
    geom_point(aes(x = `Alpha, Terminal1`, y = `Alpha, Terminal2`))
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
tab_start <-
results %>%mutate(Truth = rep(truth$truth,nsim)) %>%
group_by(Parameter) %>%
summarise(Truth = Truth[1], 
          Mean = mean(Estimate[!anyerror], trim = 0.005),
          SD = mean((Estimate[!anyerror]-Mean)^2, na.rm=T, trim = 0.005)^.5,
          Est.SD = mean(Estimate.SE[!anyerror], na.rm=T, trim = 0.005), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value)

library(gt)
gt(tab_start, rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Joint Model Frailty Pack Fits (R = 1,000, n = 1,200)**"))%>%
cols_label(
    Truth = html("True Value"),
    Mean = html("Avg. Estimate"),
    SD = html("SE"),
    Est.SD = html("Avg. Est. SE")
)
```

## Errors

```{r}
mean(results$anyerror)
mean(results$error2)
mean(results$error4)
```

# Comparison with Joint Model

```{r}
Parameter = mod.joint.a[[1]]$summary.table$Parameter
results <- list()
for(i in 1:nsim){
results[[i]]<- 
tibble(sim = i,
       Parameter = c(Parameter[c(1:4,7:8,10:11)], #model1a
	         Parameter[c(1:2,5:6,7,9,10,12)],
	         Parameter),
       Estimates = c(mod.joint.b[[i]]$b[1:5]^2, mod.joint.b[[i]]$b[6:8], 
       	         mod.joint.c[[i]]$b[1:5]^2, mod.joint.c[[i]]$b[6:8],
       	         mod.joint.a[[i]]$summary.table$Estimate) %>% round(3),
       SE = c(sqrt(diag(mod.joint.b[[i]]$varHtotal))[1:5]*2*sqrt(mod.joint.b[[i]]$b[1:5]), 
       	  sqrt(diag(mod.joint.b[[i]]$varHtotal))[6:8], 
       	  sqrt(diag(mod.joint.c[[i]]$varHtotal))[1:5]*2*sqrt(mod.joint.c[[i]]$b[1:5]),
       	  sqrt(diag(mod.joint.c[[i]]$varHtotal))[6:8],
       	  mod.joint.a[[i]]$summary.table$Estimate.SE) %>% round(3),
       Error = c(rep(mod.joint.b[[1]]$istop != 1, 8),
                 rep(mod.joint.c[[1]]$istop != 1, 8),
                 rep(mod.joint.a[[1]]$critCV[2] != 1, 12))) %>%
mutate(model = c(rep("Death",8), 
       	         rep("Discharge",8), 
       	         rep("Competing",12)))
}

results %>%
do.call(what = "rbind") %>% 
group_by(Parameter, model) %>%
summarise(Mean = median(Estimates[!Error], trim = 0.005),
          SD = sd(Estimates[!Error]),
          Est.SD = median(SE[!Error], na.rm=T), .groups = "drop"
) %>%
pivot_longer(cols = Mean:Est.SD) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value) %>%
pivot_wider(id_cols = Parameter, names_from = c(model), values_from = c(Mean, SD, Est.SD) ) %>%
#arrange(Parameter)  %>%
relocate(SD_Death, .after = Mean_Death)%>%
relocate(SD_Discharge, .after = Mean_Discharge)%>%
relocate(SD_Competing, .after = Mean_Competing)%>%
relocate(SD_Death, .after = Mean_Death)%>%
relocate(SD_Discharge, .after = Mean_Discharge)%>%
relocate(SD_Competing, .after = Mean_Competing) %>%
dplyr::select(-Est.SD_Competing:-Est.SD_Discharge) %>%
arrange(c(5:6,1:3,7:12,4))%>%
mutate(Truth = par00[c(10:12,7:9,2,4,6,1,3,5)]) %>%
gt(rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Simulation Results: Weibull Model Estimates**"))%>%
cols_label(
    Mean_Death = html("Joint: Death"),
    Mean_Discharge = html("Joint: Discharge"),
    Mean_Competing = html("Competing"),
    Mean_Death = html("Joint (Death)"),
    Mean_Discharge = html("Joint (Discharge)"),
    Mean_Competing = html("Competing"),
    SD_Death = html("(SE)"),
    SD_Discharge = html("(SE)"),
    SD_Competing = html("(SE)"),
)%>%
fmt_missing(columns = 1:8, missing_text = "")  %>%
 tab_style(
     style = list(
       cell_borders(sides = "left")
     ),
     locations = list(
       cells_body(
         columns = vars(Truth)
       )
     )
   ) %>% 
text_transform(
  locations = cells_body(columns = c(3,5,7)),
  fn = function(x) ifelse(x == "","",paste0("(",x,")"))
  )%>%
cols_align(
  align = "left",
  columns = c(3,5,7)
)
```
