---
title: "Simulation of Classic Joint Model"
output: html_document
---
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(frailtypack); library(tidyverse)
source("../delirium_package/R/simulate_competing_data.R")
source("../delirium_package/R/random_weibull.R")
source("../delirium_package/R/extract_frailtypack_coef_weibullgap.R")
```

In this version, we are going to set the starting values for the optimizer far from the true values.

# Competing Joint Model

$$
    r_{ij}(t|w_i, X_{ij}) = r_0(t) \exp(X_{i} \gamma_r + w_i) \\
    r_0(t) =  \frac{\eta_r \cdot t^{\eta_r-1}}{\beta_r^{\eta_r}} \text{(weibull hazard)}\\
    \lambda_{i}^{(1)}(t|w_i, X_{ij}) = \lambda_0^{(1)}(t) \exp(X_{i} \gamma_1 + \alpha_{1} w_i) \\
    \lambda_{i}^{(2)}(t|w_i, X_{ij}) = \lambda_0^{(2)}(t) \exp(X_{i} \gamma_2 + \alpha_{2} w_i) \\
    \lambda_0^{(1)}(t) = \frac{\eta_1 \cdot t^{\eta_1-1}}{\beta_1^{\eta_1}} \text{(weibull hazard)}\\
    \lambda_0^{(2)}(t) = \frac{\eta_2 \cdot t^{\eta_2-1}}{\beta_2^{\eta_2}} \text{(weibull hazard)}\\
    w_i \sim_{IID} N(0,\theta)
$$

$$
    \textbf{Simulation Values:}\\
    \gamma_r = 15 , \gamma_1 = 15 , \gamma_2 = 15 \text{(treatment effects)}\\
    \eta_r = 3, \eta_1 = 10, \eta_2 = 10 \text{(weibull scale)}\\
    \beta_r = 4, \beta_1 = 4, \beta_2 = 4 \text{(weibull shape, exponential)}\\
    \alpha_1 = 2;  
    \alpha_2 = -2; 
    \theta = 5 \\
    \text{Maximum Observation (Censoring) Time }=2
$$

$$
    \textbf{Starting Values:}\\
    \gamma_r = 0 , \gamma_1 = 0 , \gamma_2 = 0 \text{(treatment effects)}\\
    \eta_r = 0.5, \eta_1 = 1, \eta_2 = 1 \text{(weibull scale)}\\
    \beta_r = 1, \beta_1 = 1, \beta_2 = 1 \text{(weibull shape, exponential)}\\
    \alpha_1 = 0,
    \alpha_2 = 0,
    \theta = 0.25 
$$

Notes:

A value of 4 for the shape (beta) parameters seemed to be ok, but a value of 5 was too high.

Values of -2 and 2 for the alpha parameters were ok, but -3 and 3 were too large.

Values of 15 for the treatment effects were ok, but 20 was too large.

# Test One

```{r, fig.height = 14, fig.width=18}
set.seed(012921)
if(file.exists("multiv_model_progress.dat")) file.remove("multiv_model_progress.dat")
if(file.exists("multiv_model_parameters.dat")) file.remove("multiv_model_parameters.dat")
if(file.exists("multiv_model_inverse.dat")) file.remove("multiv_model_inverse.dat")
if(file.exists("multiv_model_derivative.dat")) file.remove("multiv_model_derivative.dat")

data <- 
simulate.competing.data(n = 500,
		truncate = 2,
		par0 = c(betaR = 1, etaR = .5,
         betaD = 1, etaD = 1,         
         betaD2 = 1, etaD2 = 1,
         theta = .5, alpha1 = 0, alpha2 = 0,
         trtR = 0, trtD = 0, trtD2 = 0))

mod0 <- 
multivPenal(
formula = Surv(t0, t, event)~cluster(id)+trt+terminal(terminal1)+terminal2(terminal2),
formula.terminalEvent=~trt,
formula.terminalEvent2=~trt,
data=data, 
initialize = FALSE, 
save.progress = T,
hazard = "Weibull",
recurrentAG=F,
maxit = 100,
init.hazard = c(10,10,10,10,10,10),
init.Theta = 3,
init.Alpha1 = 20,
init.Alpha2 = 20,
init.B = c(-5,-5,-5))

mod0$critCV
```

```{r, fig.height = 14, fig.width=18}
par <- 
read_file(file = "../package_tests/multiv_model_parameters.dat") %>%
	strsplit("\\s+") %>% unlist %>% as.numeric
df <- as.numeric(par)[-1] %>%
	matrix(ncol = 14, byrow = T) %>%
	as.data.frame

colnames(df) <- c("n","LL", "ln.betaR.5", "ln.etaR.5", "ln.betaD.5", "ln.etaD.5",
	      "ln.betaD2.5", "ln.etaD2.5", "ln.theta.5",
	      "alpha1", "alpha2",
	      "trtR", "trtD", "trtD2")
df <- df %>% as_tibble

df %>% gather("par","val",-n) %>%
ggplot() +
geom_line(aes(x = n, y = val)) +
geom_point(aes(x = n, y = val)) +
facet_wrap("par", scales = "free_y") +
theme_bw(30) 
```


# Vary Parameters (Theta Alpha)

```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 8, fig.width = 12}
set.seed(012921)
nsim = 100
par0 = rep(each = nsim, list(
       c(betaR = 1, etaR = 0.5,
         betaD = 1, etaD = 1,         
         betaD2 =1, etaD2 = 1,
         theta = 0.5, alpha1 = 0, alpha2 = 0,
         trtR = 0, trtD = 0, trtD2 = 0)
       ))
truth =
tibble(Parameter = c("betaR", "etaR",
	          "betaD", "etaD", "betaD2", "etaD2","theta","alpha1","alpha2",
		"trtR", "trtD", "trtD2"),
truth = c(betaR = 1, etaR = 0.5,
         betaD = 1, etaD = 1,         
         betaD2 =1, etaD2 = 1,
         theta = 0.5, alpha1 = 0, alpha2 = 0,
         trtR = 0, trtD = 0, trtD2 = 0), 
start2 = c(betaR = 3, etaR = 3,
         betaD = 3, etaD = 3,         
         betaD2 =3, etaD2 = 3,
         theta = 2, alpha1 = 1, alpha2 = -1,
         trtR = 2, trtD = 2, trtD2 = 2)
       )
####################################################################
# Generate Data
if(!file.exists("Test_Saves/CompetingJointModels_012621.rdata")){
data0 <- 
map(par0, 
    ~simulate.competing.data(n = 500,truncate = 2,par0 = .))

# Fit models
mod.joint.a <- list()
mod.joint.b <- list()

for(i in 1:length(par0)){
mod.joint.a[[i]] <-
multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal1) + terminal2(terminal2),
formula.terminalEvent = ~trt, 
formula.terminalEvent2= ~trt,
data = data0[[i]],
recurrentAG=FALSE,
hazard = "Weibull",
maxit = 100, initialize = F)
cat("Model ", i, "complete.\n")
}

for(i in 1:length(par0)){
mod.joint.b[[i]] <-
multivPenal(formula = Surv(t0, t, event) ~ trt + cluster(id) +
		terminal(terminal1) + terminal2(terminal2),
formula.terminalEvent = ~trt, 
formula.terminalEvent2= ~trt,
data = data0[[i]],
recurrentAG=FALSE,
hazard = "Weibull",
maxit = 100, initialize = F,
init.hazard = c(3,3,3,3,3,3),
init.Theta = 2,
init.Alpha1 = 1,
init.Alpha2 = -1,
init.B = c(2,2,2))

cat("Model ", i, "complete.\n")
}

save(data0, mod.joint.a, mod.joint.b, file = "Test_Saves/CompetingJointModels_012621.rdata")
} else load(file = "Test_Saves/CompetingJointModels_012621.rdata")

####################################################################
# Combine  Results
results <- 
tibble(data = 1:(nsim),
       par0 = par0,
       est = map(mod.joint.a, ~.$b),
       error = map(mod.joint.a, ~.$critCV[2] !=1) %>% unlist,
       est.b = map(mod.joint.b, ~.$b),
       error.b = map(mod.joint.b, ~.$critCV[2] !=1) %>% unlist) %>% 
mutate(par0 = map(par0, ~tibble(par0 = ., Parameter = names(.)))) %>%
unnest(c(est, est.b ,par0)) %>%
mutate(est = ifelse(Parameter %in% c("betaR", "etaR", "betaD","etaD","betaD2","etaD2","theta"),
	        exp(est)^2, est),
       est.b = ifelse(Parameter %in% c("betaR", "etaR", "betaD","etaD","betaD2","etaD2","theta"),
	        exp(est.b)^2, est.b))

# Plot Results
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = data, x = est, color = error)) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "red"))+
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 100, n = 500)",
     caption = "Note: For all simulations, etaR = 0.5; betaR, betaD, etaD, betaD2, etaD2 = 1; trtR = 0.5; trtD, trtD2 = 0")

# Plot Results
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = data, x = est.b, color = error.b)) +
facet_wrap("Parameter", scales = "free") +
geom_vline(data = truth, aes(xintercept = truth), color = "red") +
scale_color_manual(values = c("black", "red"))+
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 100, n = 500)",
     caption = "Note: For all simulations, etaR = 0.5; betaR, betaD, etaD, betaD2, etaD2 = 1; trtR = 0.5; trtD, trtD2 = 0")
```


```{r, message = FALSE, warning = FALSE, echo = FALSE, fig.height = 6, fig.width = 10}
results %>% 
#filter(!error) %>%
ggplot()+
geom_point(aes(y = est.b, x = est, color = error.b)) +
facet_wrap("Parameter", scales = "free") +
geom_point(data = truth, aes(x = truth, y = truth), color = "blue", shape = 3, size = 10) +
#geom_point(data = truth, aes(x = truth, y = start2), color = "red", shape = 3, size = 10) +
scale_color_manual(values = c("black", "red"))+
geom_abline(slope = 1, intercept = 0, color = "blue") +
theme_bw()+
labs(title = "Joint Model Simulation Results under Different Values of Theta, Alpha1 (R = 100, n = 500)",
     caption = "Note: For all simulations, etaR = 0.5; betaR, betaD, etaD, betaD2, etaD2 = 1; trtR = 0.5; trtD, trtD2 = 0")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
tab_start <-
results %>%
group_by(Parameter) %>%
summarise(Truth = mean(par0[!error]),
          Mean = mean(est[!error]),
          Bias = mean(est[!error] - par0[!error], na.rm=T),
          RMSE = sqrt(mean((est[!error] - par0[!error])^2, na.rm=T)),
          SD = sd(est[!error], na.rm=T),
          Bias_SD = Bias / SD,
          Failure_Percent = 100*mean(error)
) %>%
pivot_longer(cols = Mean:Failure_Percent) %>%
mutate(value = round(value, 3)) %>%
pivot_wider(names_from = c(name), values_from = value)

library(gt)
gt(data = tab_start, rowname_col = "Parameter")%>%
tab_stubhead(label = "Parameter")%>%
tab_header(
    title = md("**Joint Model Frailty Pack Fits (R = 300, n = 500)**"))%>%
cols_label(
    Truth = html("True Value"),
    Mean = html("Avg. Estimate"),
    RMSE = html("RMSE"),
    Bias = html("Bias"),
    SD = html("SE"),
    Bias_SD = html("Bias / SE"),
    Failure_Percent = html("Failure %")
)
```
