---
title: "Data Simulation and Processing for the Competing Joint Model"
author: "Lacey Etzkorn"
date: "2/5/2022"
output: html_document
---

In this appendix, we will simulate data from the competing joint model which we fit to the reduce data.
We will discretize the data so that it appears as the data made available to us from the REDUCE trial. 
Then, we will re-process the data so that it is again suitable for fitting the competing joint model.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse,quietly = T)
set.seed(654321)
```

# Parameter Values

First, we save the parameter estimates from the REDUCE results to a named vector `par0`. 
We will change the treatment effects so that treatment decreases the frequency of delirium events, 
decreases the hazard of mortality, and decreases times to discharge.
Here we also set our trial size `n` and the time of administrative censoring `K`.

```{r}
par0 <- c(shapeR = 0.34, scaleR = 6.64, shapeM = 2.59, scaleM = 78.5, shapeD = 1.51, scaleD = 8.23,
		  sigma = 0.24, alphaM = 14.7, alphaD = -6.1, betaR = -0.25, betaM = -0.25, betaD = 0.25)
n <- 1500
K <- 28
```

# Data Simulation

First, it is helpful to define a function which simulates data from a weibull distribution with a particular shape, scale, and relative hazard.

```{r}
rweibRH <- function(n, shape ,scale , rh){
	rweibull(n, shape = shape, scale = scale * rh^(-1/shape))
}
```

Now we can simulate our data set.
We define the subject id (`id`), 
treatment indicator (`trt`), 
random frailty (`w`),
latent time of death (`T1`),
latent time of discharge (`T2`),
observed endpoint (`y`),
indicators for which terminal event occurred (`terminal1` and `terminal2`),
and event times (`t`).

```{r}
df <- tibble(
id = 1:n,
trt = rbinom(n,1,.5),
	# treatment is assigned independently with equal probability
w = rnorm(n, 0, par0["sigma"]),
	# draw frailty terms from a multivariate normal
T1 = rweibRH(n,
	 shape = par0["shapeM"],
	 scale = par0["scaleM"],
	 rh = exp(w*par0["alphaM"] + par0["betaM"] * trt)),
	# draw a mortality time
T2 = rweibRH(n,
	 shape = par0["shapeD"],
	 scale = par0["scaleD"],
	 rh = exp(w*par0["alphaD"] + par0["betaD"] * trt)),
	# draw a discharge time
y = pmin(T1, T2, K),
	# determine the timing of the endpoint 
terminal1 = as.numeric(T1 < T2 & T1 < K),
terminal2 = as.numeric(T2 < T1 & T2 < K),
	# determine which endpoint we observe (death, discharge, or censoring)
t =  map2(w,trt,
          ~tibble(
          	t = rweibRH(50,
          	            shape = par0["shapeR"],
          	            scale = par0["scaleR"],
          	            rh = exp(.x + par0["betaR"] * .y)) %>%
		  	    rbind(rweibull(50, shape = 1, scale = 2)) %>%
		  	    cumsum,
		  	event = rep(1:0, length = 100)
		  	)
	)
)
	# we draw a large number (50) of iid gap times for delirium onset and 
	# alternate them with independent gap times for delirium duration and 
	# take the cumulative sum.
head(df)
```
For this data, we had to draw random variables that represent the duration of the delirium events. 
We will assume for now that they are independent of the frailty terms and follow a weibull model with
shape 1 and scale 2 (roughly two days per delirium event).

In the main paper, we did not perform this step, and simulated data based on the simplification that 
delirium events had no duration. Since we are interested in data processing here, we add this extra step.

Now, our simulated data needs further processing.
Our goal is to structure the data so that each row corresponds to a particular risk period
for a particular individual.
We will want that row to contain information about:

1. Whether the person is at risk for delirium in a given period.
2. When the period started (`tstart`)
3. When the period ended (`t`)
4. Which type of event ended the risk period (`event`, `terminal1`, or `terminal2`).

Toward this goal, we need to remove all the events that occurred after the observed terminal event. 

```{r}
df <- df %>%
dplyr::select(-T1, -T2, -w) %>%
	# throw out the unobserved variables
unnest(t) %>%
	# unnest the gap times
group_by(id) %>%
mutate(tstart = c(0, t[-n()])) %>%
	# create a variable that represents the beginning of the risk period
dplyr::filter(tstart<=y)
	# eliminate recurrent events that have risk periods which begin 
	# after the observed terminal event
head(df)
```

Now we can clean up the event indicators.
The terminal event indicators should only be 1 for the risk period 
which they terminated.
The event indicator should be 0 for the row which corresponds to the
terminal event.
The last observation time `t` needs to be set to the terminal event time.

```{r}
df <- df %>%
mutate(terminal1 = terminal1*(t > y),
	   terminal2 = terminal2*(t > y),
	   event = event*(t < y),
	   t = ifelse(t > y, y, t)) %>%
dplyr::select(-y) %>%
	#the timing of the terminal event is no longer needed as a separate variable.
ungroup
head(df)
```

We can now read the data frame and determine what happened to an individual. 
For example, we know that participant 3 received treatment. 
They experienced their first delirium event very soon (`r df$t[3]` days) after arriving at the hospital.
This delirium event ceased at `r round(df$t[4],2)` days after admission.
They were discharged (`terminal2`) `r round(df$t[5],2)` days after admission.

At this point, we could fit the competing joint model to this data. 

# Discretization

```{r}

```

# Data Processing

Data for delirium trials often comes in a wide format where each day of a participants stay (1-28) has its own column.

```{r}

```

# Model Fitting
